\documentclass[11pt, oneside]{article}
\usepackage[letterpaper, margin=2cm]{geometry}
\usepackage{MATH520}

\begin{document}
\noindent \textbf{\Large{Caleb Logemann \\
MATH 520 Methods of Applied Math II \\
Homework 1
}}

\subsection*{Section 10.9}
\begin{enumerate}
  \item[\#3] % Done
    Prove Proposition 10.1.
    Proposition 10.1 states that if $T$ is bounded on its domain then it has a
    unique norm preserving extension to $\overline{D(T)}$.
    That is to say there exists a unique linear operator
    $S:\overline{D(T)} \subset X \to Y$ such that $Sx = Tx$ for $x \in D(T)$ and
    $\norm{S} = \norm{T}$.

    \begin{proof}
      Let $S:\overline{D(T)} \subset X \to Y$ be defined as follows.
      \[
        Sx = \lim[n \to \infty]{Tx_n}
      \]
      where the sequence $\set{x_n}_{n = 1}^{\infty}$ is any sequence in $D(T)$
      that converges to $x$.
      Note that for any $x \in \overline{D(T)}$, $x$ is a limit point of $D(T)$
      so the sequence $\set{x_n}$ exists.
      Also since $T$ is bounded it is also continuous, so the limit always
      exists.

      Next I will show that $S$ is linear.
      Consider $x_1, x_2 \in \overline{D(T)}$ and $c_1, c_2 \in \CC$.
      Then there exists sequences in $D(T)$, $\set{a_n}_{n = 1}^{\infty}$ that
      converges to $x_1$ and $\set{b_n}_{n = 1}^{\infty}$ that converges to
      $x_2$.
      Now note that the sequence ${c_1 a_n + c_2 b_n}_{n = 1}^{\infty}$
      converges to $c_1 x_1 + c_2 x_2$ by the linearity of limits.
      Therefore
      \begin{align*}
        S(c_1 x_1 + c_2 x_2) &= \lim[n \to \infty]{T(c_1 a_n + c_2 b_n)}
        \intertext{Because $T$ is linear}
        S(c_1 x_1 + c_2 x_2) &= \lim[n \to \infty]{c_1 T(a_n) + c_2 T(b_n)}
        \intertext{By the linearity of limits}
        S(c_1 x_1 + c_2 x_2) &= c_1 \lim[n \to \infty]{T(a_n)} + c_2 \lim[n \to \infty]{T(b_n)} \\
        S(c_1 x_1 + c_2 x_2) &= c_1 S(x_1) + c_2 S(x_2)
      \end{align*}
      This shows that $S$ is a linear operator.

      Next I will show that $Sx = Tx$ for $x \in D(T)$.
      Let $\set{x_n}_{n = 1}^{\infty}$ converge to $x$ in $D(T)$, then because
      $T$ is continuous, $\lim[n \to \infty]{T(x_n)} = T(x)$.
      Therefore $Sx = Tx$.

      Now I will show that $\norm{S} = \norm{T}$.
      Consider the following, note that supremums are taken excluding the zero
      vector.
      \begin{align*}
        \norm{S} &= \sup*[x \in \overline{D(T)}]{\frac{\norm[Y]{Sx}}{\norm[X]{x}}} \\
         &= \sup*[x \in \overline{D(T)}]{\frac{\norm[Y]{\lim[n \to \infty]{Tx_n}}}{\norm[X]{x}}} \\
         &= \sup*[x \in \overline{D(T)}]{\frac{\lim[n \to \infty]{\norm[Y]{Tx_n}}}{\norm[X]{x}}} \\
         &\le \sup*[x \in \overline{D(T)}]{\frac{\lim[n \to \infty]{\norm{T}\norm[X]{x_n}}}{\norm[X]{x}}} \\
         &= \sup*[x \in \overline{D(T)}]{\frac{\norm{T}\norm[X]{x}}{\norm[X]{x}}} \\
         &= \sup*[x \in \overline{D(T)}]{\norm{T}} \\
         &= \norm{T} \\
      \end{align*}
      Therefore $\norm{S} \le \norm{T}$.
      Also $D(T) \subset \overline{D(T)}$, therefore
      \begin{align*}
        \norm{S} &= \sup*[x \in \overline{D(T)}]{\frac{\norm[Y]{Sx}}{\norm[X]{x}}} \\
        &\ge \sup*[x \in D(T)]{\frac{\norm[Y]{Sx}}{\norm[X]{x}}} \\
        &= \sup*[x \in D(T)]{\frac{\norm[Y]{Tx}}{\norm[X]{x}}} \\
        &= \norm{T}
      \end{align*}
      Therefore $\norm{S} \ge \norm{T}$, and both of these inequalities imply
      that $\norm{S} = \norm{T}$.

      Finally note that this extension is unique suppose that there exists some
      other extension $S'$ with the same properties.
      Consider $x \in \overline{D(T)}$, then there exists some sequence
      $\set{x_n}_{n = 1}^{\infty} \in D(T)$ that converges to $x$.
      Now consider $S'x$ since $S'$ is linear and bounded it is continuous,
      so
      \[
        S'x = \lim[n \to \infty]{S'x_n} = \lim[n \to \infty]{Tx_n} = Sx
      \]
      because $S'x = Tx$ for $x \in D(T)$.
      Therefore the extension is unique.
    \end{proof}

  \pagebreak
  \item[\#6] % Done
    Show that a linear operator $T:\CC^N \to \CC^M$ is always bounded for
    any choice of norms on $\CC^N$ and $\CC^M$.

    \begin{proof}
      Let $T:\CC^N \to \CC^M$ be a linear operator.
      It is known that any linear operator from $\CC^N \to \CC^M$ can be
      expressed as a matrix multiplication, that is there exists matrix
      $A \in \CC^{m \times n}$ such that $Tx = Ax$ for every $x \in \CC^N$.
      %It is well known that for finite dimensional vector spaces any two norms
      %are equivalently.
      %More precisely let $\norm[1]{\cdot}$ and $\norm[2]{\cdot}$ be norms on a
      %finite dimensional vector space, then there exists constants $C_1$ and $C_2$ such
      %that
      %\[
        %0 < C_1 \le \frac{\norm[1]{x}}{\norm[2]{x}} \le C_2 < \infty
      %\]
      %for any nonzero $x$ in the vector space.
      %Since both $N$ and $M$ are finite, norms on $\CC^N$ are equivalent
      %and norms on $\CC^M$ are equivalent.
      %Therefore I will let $\norm[N]{\cdot}$ represent any norm on $\CC^N$ and
      %$\norm[M]{\cdot}$ represent any norm on $\CC^M$.

      Consider the alternative definition of the norm of an operator
      \[
        \norm{T} = \sup[\norm{x} = 1]{\norm{Tx}}
      \]
      I will let $S = \set{x \in \CC^N: \norm{x} = 1}$, which is commonly known as the
      unit sphere of the vector space.
      It is well known that the unit sphere of a finite dimensional vector space
      is a compact set and this is true for any norm chosen on that space.
      It is also well known that a continous function is bounded on every compact
      set.
      Since any norm on $\CC^M$ is continuous, the value of $\norm{Ax}$ is bounded
      for any $x \in S$.
      This implies that
      \[
        \norm{T} = \sup[x \in S]{\norm{Tx}} = \sup[x \in S]{\norm{Ax}} < \infty
      \]
      Therefore the $T$ is bounded for any choice of norms on $\CC^N$ and $\CC^M$.
    \end{proof}

  \pagebreak
  \item[\#7] % Done
    If $T, T^{-1} \in \mcB(\v{H})$ show that $\p{T^*}^{-1} \in \mcB(\v{H})$ and
    $\p{T^*}^{-1} = \p{T^{-1}}^*$.

    \begin{proof}
      Let $\v{H}$ be a Hilbert Space and let $T, T^{-1} \in \mcB(\v{H})$.
      Since $T \in \mcB(\v{H})$ this implies that there exists a unique linear
      operator $T^* \in \mcB(\v{H})$ such that $\abr{Tu, v} = \abs{u, T^*v}$ for
      all $u, v \in \v{H}$.
      First I will show that $T^*$ is one-to-one so that $\p{T^*}^{-1}$ exists.
      Consider $x, y \in \v{H}$ such that $T^*x = T^*y$ or equivalently
      $T^*(x - y) = 0$.
      Now consider any $u \in \v{H}$.
      \[
        0 = \abr{u, T^*(x - y)} = \abr{Tu, x - y}
      \]
      Since $\abr{Tu, x - y} = 0$ for any $u \in \v{H}$ this implies that
      $x - y = 0$ or $x = y$.
      Therefore $T^*$ is one-to-one and $\p{T^*}^{-1}$ exists.
      Secondly I will show that $\p{T^*}^{-1} = \p{T^{-1}}^*$.
      Consider the identity operator $I$ which is defined as $Ix = x$ for all
      $x \in \v{H}$.
      Note that $T^{-1}T = I$.
      Also note that $I^* = I$ because $I \in \mcB(\v{H})$ and
      \[
        \abr{Iu, v} = \abr{u, v} = \abr{u, Iv}
      \]
      for all $u, v \in \v{H}$.
      This implies that $\p{T^{-1}T}^* = I$ or equivalently $T^* \p{T^{-1}}^* = I$
      because $T, T^{-1} \in \mcB(\v{H})$.
      Now consider any $x \in \mcB(\v{H})$
      \begin{align*}
        x &= T^* \p{T^{-1}}^* x \\
        \p{T^*}^{-1} x &= \p{T^{-1}}^* x
      \end{align*}
      This is because $\p{T^*}^{-1}$ is defined for any $x \in \v{H}$.
      Since this is true for any $x$, this implies that
      $\p{T^*}^{-1} = \p{T^{-1}}^*$.
      Now it is easy to see that since $\norm{T} = \norm{T^*}$ this implies that
      $\norm{\p{T^*}^{-1}} = \norm{T^{-1}} < \infty$.
      Also since $\p{T^*}^{-1} = \p{T^{-1}}^*$, this implies that $\p{T^*}^{-1}$
      is linear.
      Therefore $\p{T^*}^{-1} \in \mcB(\v{H})$.
    \end{proof}

  \pagebreak
  \item[\#14] % Done
    If $T \in \mcB(\v{H})$ show that $T^*$ restricted to $R(T)$ is one-to-one.

    \begin{proof}
      Let $\v{H}$ be a Hilbert space and let $T \in \mcB(\v{H})$.
      Consider $T^*$ restricted to $R(T)$.
      In other words let $S$ be a linear function such that $D(S) = R(T)$ and
      $Sx = T^*x$ for every $x \in D(S)$.
      Let $x, y \in D(S) = R(T)$ such that $Sx = Sy$.
      Since $S$ is linear this is equivalent to $S(x - y) = 0$, which implies
      that $x - y \in N(S)$.
      Note that $N(S) \subseteq N(T^*)$, so $x - y \in N(T^*)$.
      Also note that $R(T)$ is a subspace of $\v{H}$ so that $x - y \in R(T)$.
      It has already been stated that $R(T) \subset N(T^*)^{\perp}$, which
      implies that $x - y \in N(T^*)^{\perp}$.
      Since $x - y \in N(T^*)^{\perp}$ and $x - y \in N(T^*)$, $x - y$ must be
      orthogonal to itself.
      \begin{align*}
        \abr{x - y, x - y} &= 0 \\
        \norm{x - y}^2 &= 0 \\
        \norm{x - y} &= 0 \\
        x - y &= 0 \\
        x &= y
      \end{align*}
      This follows since the only element of $\v{H}$ with norm $0$ is the zero
      element.
      This shows that when $Sx = Sy$ then $x = y$.
      Therefore $S$ is one to one or $T^*$ is one to one when restricted to the
      range of $T$.
    \end{proof}
\end{enumerate}
\end{document}
